Quantifying stochastic noise in cultured\\ circadian reporter cells

### 2015-07-02 main_text.tex
addressing point on including qualitative measures

Bioluminescence rhythms at the cell culture or tissue-level are the result of the collective behavior of thousands of cells.
Transcription at the single-cell level is strongly affected by {\itshape intrinsic} cellular noise, caused by the low molecular counts of the mRNA and protein species involved. 
As a result, bioluminescence traces of individual cells are stochastic, with significant variability in both amplitude and period length from cycle to cycle \cite{Welsh2004}. 
In addition to intrinsic noise, circadian oscillations are also affected by {\itshape extrinsic} noise sources.
Extrinsic noise results from heterogeneity between cells, such as differences in size or physical environment, leading to differences on a cell-to-cell basis in expected period and amplitude.
For circadian systems, intrinsic noise has been shown to play a larger role: a single cell's variability in period from cycle-to-cycle is larger than the variability in mean period length between cells \cite{Herzog2004}.
Both sources of noise have an effect on population-level rhythms: in cell cultures that lack cell-to-cell coupling, it has been shown that  stochastic noise is manifested in damped oscillations at the population-level as individual oscillators gradually drift out of phase \cite{Nagoshi2004, Welsh2004}.
This type of behavior has also been seen in other experimental systems, such as NF-$\kappa$B signaling or yeast glycolytic oscillations \cite{Nelson2004, Aon1992}.
The amount of noise in system is therefore linked to the ability of tissue-level clocks to maintain high-amplitude rhythms.

>>>>

Bioluminescence rhythms at the cell culture or tissue-level are the result of the collective behavior of thousands of individual cells.
Transcription at the single-cell level is strongly affected by {\itshape intrinsic} cellular noise, caused by the low molecular counts of the mRNA and protein species involved. 
As a result, bioluminescence traces of individual cells are stochastic, with significant variability in both amplitude and period length from cycle to cycle \cite{Welsh2004}. 
In addition to intrinsic noise, circadian oscillations are also affected by {\itshape extrinsic} noise sources.
Extrinsic noise results from heterogeneity between cells, such as differences in size or physical environment, leading to differences on a cell-to-cell basis in expected period and amplitude.
The effects of noise in biological systems has been well-studied, and relative amounts of intrinsic and extrinsic noise can be identified through carefully designed single-cell experiments \cite{Elowitz2002}.
For circadian systems, intrinsic noise has been suspected to play a larger role: a single cell's variability in period from cycle-to-cycle is larger than the variability in mean period length between cells \cite{Herzog2004}.
However, both sources of noise have an effect on population-level rhythms: in cell cultures that lack cell-to-cell coupling, it has been shown that  stochastic noise is manifested in damped oscillations at the population-level as individual oscillators gradually drift out of phase \cite{Nagoshi2004, Welsh2004}.
This type of behavior has also been seen in other experimental systems, such as NF-$\kappa$B signaling or yeast glycolytic oscillations \cite{Nelson2004, Aon1992}.
The amount of noise in system is therefore linked to the ability of tissue-level clocks to maintain high-amplitude rhythms.
